---
title: "O2 Concentration in CA Seawater"
author: "Sachiko Lamen"
date: "1/23/2022"
output: html_document
---

```{r setup, include= TRUE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(here)
library(AICcmodavg)
library(equatiomatic)
library(kableExtra)
```

```{r}
# Read in data
seawater_samples <- read_csv(here("data", "calcofi_seawater_samples.csv"))
```

## **Introduction**
This analysis uses a subset (n = 100)  data from the CalCOFI hydrographic time-series taken off the California coast from 1949 - 2019. The purpose of the study was to understand the relationship between chemical/physical variables and O2 concentration. This report will compare linear regression models that predict O2 concentration based on several variables. AIC and 10-fold cross validation are used to determine the model that best predicts oxygen saturation.

## **Model Selection**

```{r}
# Model 1: linear regression of oxygen saturation as a function of water temperature, salinity, and phosphate concentration

f1 <- o2sat ~ t_deg_c + salinity + po4u_m
mdl1 <- lm(f1,data = seawater_samples)

# Model 2: linear regression of oxygen saturation as a function of water temperature, salinity, phosphate concentration and depth
f2 <- o2sat ~ t_deg_c + salinity + po4u_m +depth_m
mdl2 <- lm(f2, data = seawater_samples)

# Model 3: linear regression of oxygen saturation as a funciton of water temperature, salinity, phosphate concentration, depth, and chlorophyll concentration
f3 <- o2sat ~ t_deg_c + salinity + depth_m + chlor_a + po4u_m
mdl3 <- lm(f3, data = seawater_samples)

# Use AICc (because sample size is relatively small) to select the better model, push output into table 
AIC <- data.frame() %>% 
  summarize(aic_mdl1 = AICc(mdl1),
            aic_mdl2 = AICc(mdl2),
            aic_mdl3 = AICc(mdl3))

AIC %>% 
  kable(col.names = c("Model 1", "Model 2", "Model 3"), digits = 2) %>% 
  kable_styling(bootstrap_options = "striped", 
                position = "left", full_width = FALSE)

```
Model 3 (`mdl3`) looks like the better model: ∆AIC is >2 for both model 1 model 2, suggesting that model 3 is significantly better than the other two models. This is supported by the AICcWt of 0.92 for model 3.

## **K-fold Cross-Validation**
```{r}
# Run 10-fold cross-validation on both models
folds <- 10 
fold_vec <- rep(1:folds, length.out = nrow(seawater_samples)) 
table(fold_vec)

# Set a seed number for random number generation
set.seed(13)

# Add new column named 'group' that contains the samples from the fold_vec that has a sample size of the data frame (100)
seawater_fold <- seawater_samples %>%
  mutate(group = sample(fold_vec, size = n(), replace = FALSE))

# Create RMSE function in order to score models
calc_rmse <- function(x,y) {rmse_result <- (x - y)^2 %>% mean() %>% sqrt()
return(rmse_result)}

## Calculate RMSE over all folds and take the average (using a for-loop)
# Set up blank data frame
rmse_df <- data.frame()

# Run for-loop
for(i in 1:folds) {
  kfold_test_df <- seawater_fold %>%
    filter(group == i)
  kfold_train_df <- seawater_fold %>%
    filter(group != i)
  
  kfold_mdl1 <- lm(f1, data = kfold_train_df)
  kfold_mdl2 <- lm(f2, data = kfold_train_df)
  kfold_mdl3 <- lm(f3, data = kfold_train_df)
  
  kfold_pred_df <- kfold_test_df %>%
    mutate(mdl1 = predict(kfold_mdl1, kfold_test_df),
           mdl2 = predict(kfold_mdl2, .),
           mdl3 = predict(kfold_mdl3,.))
  kfold_rmse <- kfold_pred_df %>%
    summarize(rmse_mdl1 = calc_rmse(mdl1, o2sat),
              rmse_mdl2 = calc_rmse(mdl2, o2sat),
              rmse_mdl3 = calc_rmse(mdl3, o2sat))
  
  rmse_df <- bind_rows(rmse_df, kfold_rmse) # this will add new row on bottom of data frame each time through the loop
}

rmse_summary <- rmse_df %>%
  summarise(mean_rmse_mdl1 = mean(rmse_mdl1),
            mean_rmse_mdl2 = mean(rmse_mdl2),
            mean_rmse_mdl3 = mean(rmse_mdl3))

rmse_summary %>%
    kable(col.names = c("Model 1", "Model 2", "Model 3"), digits = 3) %>% 
  kable_styling(bootstrap_options = "striped", 
                position = "left", full_width = FALSE)
```
After running a 10-fold cross validation, we find that model 3 is actually the worst at predicting o2sat and model 2 is th best (mean rmse for mdl2 = 4.86) This means that model 2 most accurately predicts the *actual* observed values for [O2] compared to the other models. 

```{r}
# Train final model (`final_mdl`) based on `f2` formula
final_mdl <- lm(f2, data = seawater_samples)

```
Final Model: `r equatiomatic::extract_eq(final_mdl, wrap = TRUE, use_coefs = TRUE)`

## **Citation**
Data citation: CalCOFI data are available for use without restriction. Data downloaded from https://calcofi.org/ccdata.html.  Accessed 1/23/2022.







